---
title: "Linear_Regression_Package_Group3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear_Regression_Package_Group3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
Heya guys! Are you looking for some Regression Analysis?
Then stick with us right here!

Package name: **Linear_Regression_Package_Group3 (yes, no creativity!)**

GitHub repository: [Group3 Github Link](https://github.com/AU-R-Programming/Final-Project-Group3)

Published GitHub repository: [Group3 Github Link](https://github.com/AU-R-Programming/Final-Project-Group3)


Our package is the implementation of the main basic functions that one would like to play around with Linear Regression. Yes! plots of residuals, F-test statistics, computation of the mean square prediction error.. P-value? Of course!.

This is a group authorship. lol, talking about group, here are the authors:

**John M. Musah**

**Alan Galloway**

**Colin Clifton**



## Linear Regression

The following package is uploaded to GitHub publicly and is available for download in the AU-R-Programming repository with our repo being named:  <br />  Final-Project-Group3.

To install and use the package you can do the following.

```{r message=FALSE, warning=FALSE}
devtools::install_github('AU-R-Programming/Final-Project-Group3/Group3Package/FinalProjPackage')
library(FinalProjPackage)
```

## Data Selection

for data you are able to use any data you would like to perform linear regression on. For simplicity we will demo with data already included with R
```{r message=FALSE, warning=FALSE}
data("USArrests")
```

## Estimating Beta
The first function we implemented was the for estimating $\hat{\beta}$ in the linear regression model. The parameters needed for this function is a dependent variable of interest with vector length n, and a dependent variable with a matrix dimension of (n x p) with the first column containing only 1's.


```{r}
coeff_beta <- function (Dependent, Independent) {
  
  Dependent <- as.vector(Dependent)
  #cbind to add a column of ones at the front.
  Independent <- as.matrix(cbind(1,Independent))
  
  beta <- solve(t(Independent)%*%Independent)%*%t(Independent)%*%Dependent
  
  return(beta)
}
```

The output of this function using the USAressts data with our Dependent variable being ('USArrests\$Murder') and our Independent variable being ('USArrests\$UrbanPop') is the following.

```{r, echo=FALSE}
coeff_beta(USArrests$Murder,USArrests$UrbanPop)
```
This value allows us to compute the rest of our functions in the package.

## Confidence Intervals

The next function in our package we will introduce is for calculating confidence intervals. Similarly this function requires an independent and dependent variable. In addition, we allow the user to select a value for alpha however this is not required as the default is set at .05. Here we calculate a $\sigma^2$ value in order to get a correct value for the interval.

```{r}
coeff_ci <- function (Dependent, Independent, alpha = 0.05) {
  
  beta <- coeff_beta(USArrests$Murder, USArrests$UrbanPop)
  b <- beta
  Dependent <- as.vector(USArrests$Murder)
  Independent <- as.matrix(USArrests$UrbanPop)
  n <- length(Dependent)
  p <- dim(Independent)[2]
  degrees_freedom <- n - p
  Independent <- as.matrix(cbind(1,USArrests$UrbanPop))
  
  #resid
  resid <- Dependent - Independent%*%as.matrix(b)
  
  #sigma2.hat
  sigma2.hat <- (1/degrees_freedom) * t(resid) %*% resid
  
  #var.beta
  var.beta <- as.numeric(sigma2.hat)*solve(t(Independent)%*%Independent)
  
  #quant
  quant <- 1 - alpha/2
  
  a <- beta - qnorm(p = quant)
  a <- as.numeric(a)
  
  c <- sqrt(var.beta)
  c <- t(b)
  c <- as.numeric(b)
  
  d <- beta + qnorm(p = quant)
  d <- as.matrix(d)
  d <- as.numeric(d)
  
  ci.beta <- c(a%*%c, d%*%c)
  
  return(ci.beta)
}
```

We use the same data as we did before with the default alpha value and we get the following:

```{r echo=FALSE, warning=FALSE}
coeff_ci(USArrests$Murder,USArrests$UrbanPop)
```


## $R^2$ calcuation

The next function in our package we would will introduce is the $R^2$ function. Similarly this function requires an independent and dependent variable. the $R^2$ function shows us how well our linear model fits the data.

```{r}
coeff_rsquared <- function (Dependent, Independent) {

  beta <- coeff_beta(Dependent, Independent)
  b <- beta
  Dependent <- as.vector(Dependent)
  Independent <- as.matrix(Independent)
  n <- length(Dependent)
  p <- dim(Independent)[2]
  degrees_freedom <- n - p
  sample_mean <- mean(Dependent)
  Ind <- cbind(1,Independent)
  y_hat <- Ind%*% as.matrix(b)
  fit <- Ind%*%as.matrix(b)
  
  SSE <- sum((Dependent - y_hat)^2)
  SST <- sum((Dependent - sample_mean)^2)

  r_squared <- 1 - SSE/SST
  
  return(r_squared)
}
```

The output with the same Independent and Dependent variables is as follows:
```{r echo=FALSE, warning=FALSE}
coeff_rsquared(USArrests$Murder,USArrests$UrbanPop)
```



## Mallows' $Cp$ calcuation

Mallows' $Cp$ is another function for estimating how well the linear model fits the data. Similarly this function requires an independent and dependent variable. The function is as follows:
```{r}
coeff_mallows_Cp <- function (Dependent, Independent) {
  
  beta <- coeff_beta(Dependent, Independent)
  b <- beta
  Dependent <- as.vector(Dependent)
  Independent <- as.matrix(Independent)
  n <- length(Dependent)
  p <- dim(Independent)[2]
  degrees_freedom <- n - p
  sample_mean <- mean(Dependent)
  Ind <- cbind(1,Independent)
  y_hat <- Ind%*% as.matrix(b)
  fit <- Ind%*%as.matrix(b)
  
  SSE <- sum((Dependent - y_hat)^2)
  
  #resid
  resid <- Dependent - Ind%*%as.matrix(b)
  
  #sigma2.hat
  sigma2.hat <- (1/degrees_freedom) * t(resid) %*% resid
  
  mallows <- SSE +2*p*sigma2.hat
  
  return(mallows)
  
}
```

With outputs from out data being the following:

```{r echo=FALSE, warning=FALSE}
coeff_mallows_Cp(USArrests$Murder,USArrests$UrbanPop)
```


## F-test
The F-test function is a test to determine whether the model is explaining enough in comparison to a simple average. This test returns 4 values. First, is the F-statistic denoted as $F^*$. Second is the Degrees of Freedom. Third is the Degrees of Freedom for Error. With the final value being our p-value being $P(F>F^*)$.
```{r}
coeff_ftest <- function (Dependent, Independent) {

  beta <- coeff_beta(Dependent, Independent)
  b <- beta
  Dependent <- as.vector(Dependent)
  Independent <- as.matrix(Independent)
  n <- length(Dependent)
  sample_mean <- mean(Dependent)
  Ind <- cbind(1,Independent)
  y_hat <- Ind%*% as.matrix(b)
  fit <- Ind%*%as.matrix(b)
  p <- dim(Ind)[2]
  degrees_freedom <- n - p
  
  DFM <- p - 1
  DFE <- n - p
  SSM <- sum((y_hat - sample_mean)^2)
  
  MSM <- SSM/DFM
  
  SSE <- sum((Dependent - y_hat)^2)
  
  MSE <- SSE/DFE
  
  f_stat <- MSM/MSE
 
  f_prob <- pf(f_stat, DFM, DFE, lower.tail = FALSE)
  f_info <- rbind(f_stat, DFM, DFE, f_prob)
  rownames(f_info) <- c("F Stat: ","Lower DF: ","Higher DF:","P-Value: ")
  return(f_info) 
  
}
```
The F-Test function returns the following with our data.
```{r echo=FALSE, warning=FALSE}
coeff_ftest(USArrests$Murder,USArrests$UrbanPop)
```



## Plot Creation

The following function creates our 3 plots. The first is a residuals vs fitted-values. The second is qq-plot of residuals. The third is a Histogram of residuals.
```{r}
coeff_plot <- function (Dependent, Independent) {
 
  beta <- coeff_beta(Dependent, Independent)
  Dependent <- as.matrix(Dependent)
  Independent <- as.matrix(cbind(1,Independent))
  
  n <- length(Dependent)
  p <- dim(Independent)[2]
  degrees_freedom <- n - p
  ci_beta <- rep(NA, p)
  size <- 1000
  residual <- Dependent - Independent%*%as.matrix(beta)
  fit <- Independent%*%beta

  par(mfrow=c(1,3),oma = c(1,0,1,0))
  plot(residual~fit, xlab="Residual Values",ylab="Fitted Values",main="Residuals vs Fitted")
  qqnorm(residual)
  HistResVFit=hist(residual, xlab = "Residual Values", ylab="Frequency")
  title(line = -2)

}
```

Using our data the plots return the following.

```{r echo=FALSE, warning=FALSE}
coeff_ftest(USArrests$Murder,USArrests$UrbanPop)
```


In addition to these functions we included help documentation for each of them this can be viewed from your machine by running ?'function_name'. Below is how you would access the help documentation for each function.
```{r}
?coeff_beta
?coeff_ci
?coeff_rsquared
?coeff_mallows_Cp
?coeff_ftest
?coeff_plot
```

## References

https://economictheoryblog.com/2016/02/20/rebuild-ols-estimator-manually-in-r/
<br />
https://github.com/andrewc94/Linear-Regression-in-R
<br />
https://stackoverflow.com/questions/40901445/function-to-calculate-r2-r-squared-in-r
<br />
https://smac-group.github.io/ds/section-functions.html#section-example-continued-least-squares-function
<br />
https://smac-group.github.io/ds/section-r-packages.html
